---
title: "Portfolio Management and Risk Consultation"
subtitle: "FINSTAD Case 2 -- Group 3 (Charlie)"
author:
  - "CRUZ, Ricardo Miguel Iñigo"
  - "GALEDO, Enrique Lorenzo Hermoso"
  - "SEBALLOS, Josiah Dweyn Panganiban"
  - "SEECHUNG, Camille Castro"
  - "SISON, Aaron Joshua Estacio"
date: "August 17, 2026"
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 3
    number-sections: true
    colorlinks: true
    documentclass: article
    classoption: 12pt
    geometry:
      - margin=1in
    fontsize: 11pt
    code-block-font-size: small
    highlight-style: arrow
    keep-tex: true
include-in-header:
  - text: |
      \usepackage{fvextra}
      \fvset{breaklines=true,breakanywhere=true}
      \RecustomVerbatimEnvironment{Highlighting}{Verbatim}{breaklines=true,breakanywhere=true,commandchars=\\\{\}}
      \let\verbatim\Verbatim
      \let\endverbatim\endVerbatim
      \usepackage{microtype}
      \usepackage{adjustbox}
      \usepackage{booktabs}
      \usepackage{tabularx}
      \usepackage{pdflscape}
      \usepackage{float}
      \usepackage{ragged2e}
      \usepackage{csquotes}
      \clubpenalty=10000
      \widowpenalty=10000
      \displaywidowpenalty=10000
      \sloppy
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 4.8,
  dpi = 300,
  fig.pos = "H"
)

# Safe kable helper for PDF and HTML rendering
safe_kable <- function(data, caption = NULL, digits = 4) {
  knitr::kable(data, caption = caption, digits = digits, booktabs = TRUE) %>%
    kableExtra::kable_styling(
      font_size = 8,
      latex_options = c("scale_down", "HOLD_position", "striped"),
      bootstrap_options = c("striped", "hover", "condensed")
    )
}
```

# Setup and Library Loading

Load the quantitative finance, portfolio analytics, and data visualization libraries required for the analysis.

```{r libraries}
# Core data manipulation and visualization
library(tidyverse)
library(lubridate)
library(scales)

# Time series and quantitative financial analysis
library(xts)
library(zoo)
library(PerformanceAnalytics)
library(PortfolioAnalytics)

# Solvers for portfolio optimization
library(ROI)
library(ROI.plugin.quadprog)

# Table formatting and visual aesthetics
library(corrplot)
library(knitr)
library(kableExtra)
```

# Data Ingestion and Preparation

Load the compiled master dataset containing historical daily close prices for Group 3's asset universe (TEL, MER, AEV, NVDA, META, BTC, SPY) and the risk-free rate benchmark over the sample period (January 2, 2020 -- December 31, 2025).

```{r data-ingestion}
# Define path to master dataset
data_path <- file.path("..", "data", "master_dataset.csv")

if (!file.exists(data_path)) {
  data_path <- file.path("data", "master_dataset.csv")
}

raw_data <- read_csv(data_path, show_col_types = FALSE)

# Preview structure
glimpse(raw_data)
```

## Daily Log Return Computation

Compute daily log returns for all portfolio assets and the benchmark:

$$r_{i,t} = \ln\left(\frac{P_{i,t}}{P_{i,t-1}}\right)$$

```{r return-computation}
# Sort chronologically by date
prices_df <- raw_data %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date)

# Extract asset price columns
asset_cols <- c("TEL", "MER", "AEV", "NVDA", "META", "BTC", "SPY")

# Convert matrix to xts time series object
prices_mat <- as.matrix(prices_df[, asset_cols])
prices_xts <- xts(prices_mat, order.by = prices_df[["Date"]])

# Compute daily log returns
returns_xts <- Return.calculate(prices_xts, method = "log")[-1, ]

# Separate portfolio assets from benchmark
portfolio_assets <- c("TEL", "MER", "AEV", "NVDA", "META", "BTC")
benchmark_asset <- "SPY"

asset_returns_xts <- returns_xts[, portfolio_assets]
benchmark_returns_xts <- returns_xts[, benchmark_asset]

# Risk-free rate (convert annual T-Bill rate to daily log-equivalent rate)
rf_annual <- mean(prices_df[["RF_Rate"]], na.rm = TRUE)
rf_daily <- log(1 + rf_annual) / 252

cat("Dataset span:", format(min(index(returns_xts))), "to", format(max(index(returns_xts))), "\n")
cat("Total trading days:", nrow(returns_xts), "\n")
cat("Annualized Risk-Free Rate:", percent(rf_annual, accuracy = 0.01), "\n")
cat("Daily Log Risk-Free Rate:", percent(rf_daily, accuracy = 0.0001), "\n")
```

## Data Validation and Summary Check

Verify data integrity, check for missing values, and print head records of returns.

```{r}
#| label: tbl-data-validation
#| tbl-cap: "First 6 Days of Asset Daily Log Returns (2020)"

# Verify no missing values in returns time series
missing_summary <- colSums(is.na(returns_xts))
stopifnot(all(missing_summary == 0))

# Format first 6 rows for table display
returns_head_df <- as.data.frame(head(returns_xts)) %>%
  rownames_to_column(var = "Date") %>%
  mutate(across(-Date, ~ sprintf("%.4f", .x)))

kbl(returns_head_df, booktabs = TRUE, align = "c") %>%
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = c("striped", "hover"))
```

# Exploratory Financial Analysis

```{r}
#| label: tbl-exploratory-stats
#| tbl-cap: "Descriptive Statistics for Portfolio Assets (2020--2025)"

# Compute descriptive statistics
stats_mat <- table.Stats(asset_returns_xts)

# Convert to data frame for clean rendering
stats_df <- as.data.frame(stats_mat) %>%
  rownames_to_column(var = "Metric")

kbl(stats_df, booktabs = TRUE, digits = 4) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), bootstrap_options = c("striped", "hover"))
```

## Quantitative Design System (QVRS Theme & Colors)

```{r quant-theme}
# QVRS Economist Color Tokens
qvrs_colors <- c(
  "TEL"  = "#1F2E7A", # Chicago 30
  "MER"  = "#2E45B8", # Chicago 45
  "AEV"  = "#475ED1", # Chicago 55
  "NVDA" = "#1DC9A4", # Hong Kong 45
  "META" = "#F97A1F", # Singapore 55
  "BTC"  = "#E3120B", # Economist Red
  "SPY"  = "#595959"  # London 35
)

# Custom Economist/QVRS ggplot2 theme
theme_quant <- function() {
  theme_minimal() +
    theme(
      text = element_text(color = "#0D0D0D"),
      plot.title = element_text(face = "bold", size = 12, margin = margin(b = 4)),
      plot.subtitle = element_text(size = 9.5, color = "#333333", margin = margin(b = 8)),
      plot.caption = element_text(size = 8, color = "#595959", hjust = 0, margin = margin(t = 6)),
      axis.title = element_text(size = 9, face = "bold", color = "#1A1A1A"),
      axis.text = element_text(size = 8, color = "#333333"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#F2F2F2", linewidth = 0.3),
      legend.position = "bottom",
      legend.title = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_rect(fill = "#EBEDFA", color = NA),
      strip.text = element_text(face = "bold", size = 9, color = "#141F52")
    )
}
```

## Historical Price & Return Trajectories

```{r}
#| label: fig-ts-returns
#| fig.cap: "Daily Log Return Trajectories Across Portfolio Universe (2020--2025)"
#| fig.width: 8
#| fig.height: 6

# Reshape returns for tidy plotting
returns_tidy <- as.data.frame(returns_xts) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Asset", values_to = "Return") %>%
  mutate(Asset = factor(Asset, levels = names(qvrs_colors)))

ggplot(returns_tidy, aes(x = Date, y = Return, color = Asset)) +
  geom_line(linewidth = 0.35, alpha = 0.85) +
  facet_wrap(~ Asset, ncol = 2, scales = "free_y") +
  scale_color_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_date(date_labels = "%Y", date_breaks = "2 years") +
  labs(
    title = "Daily Log Return Dynamics and Volatility Clustering",
    subtitle = "Sample Period: January 2, 2020 through December 31, 2025",
    x = NULL,
    y = "Daily Log Return"
  ) +
  theme_quant() +
  theme(legend.position = "none")
```

## Distributional Shape: Histograms and Empirical Density vs Normal Reference

```{r}
#| label: fig-density-hist
#| fig.cap: "Empirical Return Distributions vs Gaussian Reference Curves"
#| fig.width: 8
#| fig.height: 6

# Compute normal distribution parameters per asset for reference curves
norm_params <- returns_tidy %>%
  group_by(Asset) %>%
  summarise(
    mean = mean(Return, na.rm = TRUE),
    sd = sd(Return, na.rm = TRUE),
    .groups = "drop"
  )

# Grid of x values for theoretical normal density overlay
grid_df <- norm_params %>%
  group_by(Asset) %>%
  reframe(
    x = seq(mean - 4 * sd, mean + 4 * sd, length.out = 200),
    y = dnorm(x, mean = mean, sd = sd)
  )

ggplot(returns_tidy, aes(x = Return)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 60,
    fill = "#D6DBF5",
    color = "#1F2E7A",
    linewidth = 0.2,
    alpha = 0.7
  ) +
  geom_density(aes(color = "Empirical Density"), linewidth = 0.8) +
  geom_line(
    data = grid_df,
    aes(x = x, y = y, color = "Gaussian Normal"),
    linetype = "dashed",
    linewidth = 0.75
  ) +
  facet_wrap(~ Asset, scales = "free", ncol = 3) +
  scale_color_manual(
    values = c("Empirical Density" = "#141F52", "Gaussian Normal" = "#E3120B")
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Empirical Distribution Shape: Heavy Tails and Non-Normality",
    subtitle = "Comparison of empirical kernel density against theoretical normal distribution",
    x = "Daily Log Return",
    y = "Density"
  ) +
  theme_quant()
```

## Quantile-Quantile (Q-Q) Tail Diagnostics

```{r}
#| label: fig-qq-plots
#| fig.cap: "Normal Q-Q Diagnostics Demonstrating Heavy Tail Leptokurtosis"
#| fig.width: 8
#| fig.height: 5.5

ggplot(returns_tidy, aes(sample = Return)) +
  stat_qq_line(color = "#E3120B", linewidth = 0.7, linetype = "dashed") +
  stat_qq(color = "#1F2E7A", alpha = 0.5, size = 0.8) +
  facet_wrap(~ Asset, scales = "free", ncol = 3) +
  labs(
    title = "Quantile-Quantile Tail Departure from Normality",
    subtitle = "S-shaped curves confirm pronounced fat-tail behavior across asset returns",
    x = "Theoretical Gaussian Quantiles",
    y = "Sample Quantiles"
  ) +
  theme_quant()
```

## Return Dispersion & Extreme Outlier Comparison

```{r}
#| label: fig-boxplots
#| fig.cap: "Return Dispersion and Extreme Outlier Boxplots"
#| fig.width: 8
#| fig.height: 4.5

ggplot(returns_tidy, aes(x = Asset, y = Return, fill = Asset)) +
  geom_boxplot(outlier.color = "#E3120B", outlier.size = 1, outlier.alpha = 0.5, linewidth = 0.4) +
  scale_fill_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Cross-Asset Return Dispersion and Extreme Tail Outliers",
    subtitle = "Comparison of median, interquartile range (IQR), and extreme shocks",
    x = "Asset",
    y = "Daily Log Return"
  ) +
  theme_quant() +
  theme(legend.position = "none")
```

## Cumulative Performance Comparison (Growth of $1.00)

```{r}
#| label: fig-cum-returns
#| fig.cap: "Cumulative Growth Trajectories of USD 1.00 Investment (2020--2025)"
#| fig.width: 8
#| fig.height: 4.8

# Compute cumulative wealth index
cum_returns_df <- as.data.frame(cumprod(1 + returns_xts)) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Asset", values_to = "Wealth") %>%
  mutate(Asset = factor(Asset, levels = names(qvrs_colors)))

ggplot(cum_returns_df, aes(x = Date, y = Wealth, color = Asset)) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = qvrs_colors) +
  scale_y_log10(labels = label_dollar(prefix = "USD ")) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  labs(
    title = "Cumulative Wealth Index Growth Trajectories (Log Scale)",
    subtitle = "Hypothetical USD 1.00 investment performance from January 2020 to December 2025",
    x = NULL,
    y = "Wealth Index (USD)"
  ) +
  theme_quant()
```

## Correlation Structure

```{r}
#| label: fig-correlation-matrix
#| fig.cap: "Asset Return Correlation Heatmap (Group 3 Universe)"

# Compute pairwise correlation matrix
cor_matrix <- cor(asset_returns_xts, use = "pairwise.complete.obs")

# Render correlation heatmap
corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.col = "black",
  number.cex = 0.8,
  mar = c(0, 0, 1, 0)
)
```

# Portfolio Construction (Equal-Weight vs Monthly Rebalanced)

This section evaluates the performance of equal-weighted asset allocations under two distinct execution regimes: a static Buy-and-Hold strategy and a disciplined Monthly Rebalanced strategy. The 6-asset portfolio universe (TEL, MER, AEV, NVDA, META, BTC) starts with equal baseline weights ($w_i = 1/6 \approx 16.67\%$) on January 2, 2020 and is benchmarked against the SPDR S&P 500 ETF Trust (SPY).

```{r portfolio-construction-calc}
# Equal weight allocation vector across the 6 portfolio assets
w_eq <- rep(1 / 6, 6)
names(w_eq) <- colnames(asset_returns_xts)

# Strategy 1: Buy-and-Hold Equal-Weighted Portfolio
port_bnh <- Return.portfolio(
  R = asset_returns_xts,
  weights = w_eq,
  verbose = TRUE
)
r_bnh <- port_bnh$returns
weights_bnh <- port_bnh$EOP.Weight

# Strategy 2: Monthly Rebalanced Equal-Weighted Portfolio
port_reb <- Return.portfolio(
  R = asset_returns_xts,
  weights = w_eq,
  rebalance_on = "months",
  verbose = TRUE
)
r_reb <- port_reb$returns
weights_reb <- port_reb$EOP.Weight

# Risk-free rate metrics (5.25% per annum)
rf_annual <- 0.0525
rf_daily <- log(1 + rf_annual) / 252

# Performance evaluation helper function
eval_portfolio <- function(r_series, name) {
  cum_ret <- exp(sum(r_series)) - 1
  ann_ret <- as.numeric(Return.annualized(r_series, geometric = TRUE))
  ann_sd  <- as.numeric(StdDev.annualized(r_series))
  sharpe  <- (ann_ret - rf_annual) / ann_sd
  max_dd  <- as.numeric(maxDrawdown(r_series))
  
  data.frame(
    Strategy = name,
    `Cumulative Return` = sprintf("%.2f%%", cum_ret * 100),
    `CAGR (Annualized)` = sprintf("%.2f%%", ann_ret * 100),
    `Annual Volatility` = sprintf("%.2f%%", ann_sd * 100),
    `Sharpe Ratio (Rf=5.25%)` = sprintf("%.4f", sharpe),
    `Max Drawdown` = sprintf("%.2f%%", max_dd * 100),
    check.names = FALSE
  )
}

perf_summary_df <- rbind(
  eval_portfolio(r_bnh, "Equal-Weight (Buy & Hold)"),
  eval_portfolio(r_reb, "Equal-Weight (Monthly Rebalanced)"),
  eval_portfolio(returns_xts[, "SPY"], "SPDR S&P 500 ETF (SPY Benchmark)")
)
```

## Comparative Performance Evaluation

```{r}
#| label: tbl-portfolio-performance
#| tbl-cap: "Portfolio Performance Summary Comparison (2020--2025)"

safe_kable(perf_summary_df, caption = "Portfolio Performance Summary Comparison (2020--2025)")
```

## Cumulative Portfolio Growth Comparison

```{r}
#| label: fig-portfolio-growth
#| fig.cap: "Cumulative Growth Trajectories of USD 1.00 Investment Across Strategies"
#| fig.width: 8
#| fig.height: 5

# Merge portfolio return series
port_returns_combined <- cbind(r_bnh, r_reb, returns_xts[, "SPY"])
colnames(port_returns_combined) <- c("Buy_Hold", "Monthly_Rebalanced", "SPY_Benchmark")

# Compute cumulative wealth index
port_growth_df <- as.data.frame(cumprod(1 + port_returns_combined)) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Strategy", values_to = "Wealth") %>%
  mutate(
    Strategy = case_when(
      Strategy == "Buy_Hold" ~ "Equal-Weight (Buy & Hold)",
      Strategy == "Monthly_Rebalanced" ~ "Equal-Weight (Monthly Rebalanced)",
      Strategy == "SPY_Benchmark" ~ "SPDR S&P 500 ETF (SPY)"
    ),
    Strategy = factor(
      Strategy,
      levels = c("Equal-Weight (Buy & Hold)", "Equal-Weight (Monthly Rebalanced)", "SPDR S&P 500 ETF (SPY)")
    )
  )

strat_colors <- c(
  "Equal-Weight (Buy & Hold)" = "#E3120B",        # Economist Red
  "Equal-Weight (Monthly Rebalanced)" = "#1F2E7A", # Chicago 30
  "SPDR S&P 500 ETF (SPY)" = "#595959"            # London 35
)

ggplot(port_growth_df, aes(x = Date, y = Wealth, color = Strategy, linetype = Strategy)) +
  geom_line(linewidth = 0.85) +
  scale_color_manual(values = strat_colors) +
  scale_linetype_manual(values = c("solid", "solid", "dashed")) +
  scale_y_log10(labels = label_dollar(prefix = "USD ")) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  labs(
    title = "Cumulative Wealth Growth Comparison (Log Scale)",
    subtitle = "Performance of USD 1.00 allocation from January 2020 through December 2025",
    x = NULL,
    y = "Wealth Index (USD)"
  ) +
  theme_quant()
```

## Portfolio Asset Weight Drift Analysis

```{r}
#| label: fig-rebalanced-weights
#| fig.cap: "Asset Weight Drift Over Time in the Un-Rebalanced Buy-and-Hold Portfolio"
#| fig.width: 8
#| fig.height: 5.2

# Reshape end-of-period weights for Buy & Hold strategy
weights_bnh_df <- as.data.frame(weights_bnh) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Asset", values_to = "Weight") %>%
  mutate(Asset = factor(Asset, levels = names(qvrs_colors)))

ggplot(weights_bnh_df, aes(x = Date, y = Weight, fill = Asset)) +
  geom_area(alpha = 0.85, color = "#FFFFFF", linewidth = 0.1) +
  scale_fill_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  labs(
    title = "Asset Allocation Drift in Buy-and-Hold Execution",
    subtitle = "Evolution of asset weights driven by differential market performance (2020--2025)",
    x = NULL,
    y = "Portfolio Weight Share"
  ) +
  theme_quant()
```

# Portfolio Optimization (Efficient Frontier, GMV, Max Sharpe)

```{r portfolio-optimization}
# Placeholder for Portfolio Optimization routines
```

# Portfolio Risk Assessment (VaR, CVaR, Monte Carlo, Stress Testing)

```{r portfolio-risk}
# Placeholder for Portfolio Risk Assessment routines
```

# Executive Investment Recommendation

```{r executive-recommendation}
# Placeholder for final quantitative recommendation summary
```
