---
title: "Portfolio Management and Risk Consultation"
subtitle: "FINSTAD Case 2 -- Group 3 (Charlie)"
author: "Group 3: Ricardo Cruz, Enrique Galedo, Josiah Seballos, Camille Seechung, Aaron Sison"
date: "August 17, 2026"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    code-summary: "Show R Code"
---

```{r setup, include=FALSE}
knitr::opts_chunk_set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 5,
  dpi = 300
)
```

# Setup and Library Loading

Load the quantitative finance, portfolio analytics, and data visualization libraries required for the analysis.

```{r libraries}
# Core data manipulation and visualization
library(tidyverse)
library(lubridate)
library(scales)

# Time series and quantitative financial analysis
library(xts)
library(zoo)
library(PerformanceAnalytics)
library(PortfolioAnalytics)

# Solvers for portfolio optimization
library(ROI)
library(ROI.plugin.quadprog)
library(ROI.plugin.glpk)

# Table formatting and visual aesthetics
library(corrplot)
library(knitr)
library(kableExtra)
```

# Data Ingestion and Preparation

Load the compiled master dataset containing historical daily close prices for Group 3's asset universe (TEL, MER, AEV, NVDA, META, BTC, SPY) and the risk-free rate benchmark over the sample period (January 2, 2020 -- December 31, 2025).

```{r data-ingestion}
# Define path to master dataset
data_path <- file.path("..", "data", "master_dataset.csv")

# Read master dataset
if (!file.exists(data_path)) {
  data_path <- file.path("data", "master_dataset.csv")
}

raw_data <- read_csv(data_path)

# Preview structure
glimpse(raw_data)
```

## Daily Log Return Computation

Compute daily log returns for all portfolio assets and the benchmark:

$$r_{i,t} = \ln\left(\frac{P_{i,t}}{P_{i,t-1}}\right)$$

```{r return-computation}
# Sort by date
prices_df <- raw_data %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date)

# Extract asset price columns
asset_cols <- c("TEL", "MER", "AEV", "NVDA", "META", "BTC", "SPY")

# Convert to xts object for quantitative analysis
prices_xts <- xts(prices_df[, asset_cols], order.by = prices_df$Date)

# Compute daily log returns
returns_xts <- Return.calculate(prices_xts, method = "log")[-1, ]

# Separate portfolio assets from benchmark
portfolio_assets <- c("TEL", "MER", "AEV", "NVDA", "META", "BTC")
benchmark_asset <- "SPY"

asset_returns_xts <- returns_xts[, portfolio_assets]
benchmark_returns_xts <- returns_xts[, benchmark_asset]

# Risk-free rate (annualized PH 91-Day T-Bill average rate converted to daily rate)
rf_annual <- mean(prices_df$RF_Rate, na.rm = TRUE)
rf_daily <- (1 + rf_annual)^(1/252) - 1

cat("Dataset span:", format(min(index(returns_xts))), "to", format(max(index(returns_xts))), "\n")
cat("Total trading days:", nrow(returns_xts), "\n")
cat("Annualized Risk-Free Rate:", percent(rf_annual, accuracy = 0.01), "\n")
cat("Daily Risk-Free Rate:", percent(rf_daily, accuracy = 0.0001), "\n")
```

## Data Validation and Summary Check

Verify data integrity, check for missing values, and print head and tail records of returns.

```{r data-validation}
# Missing value check
missing_summary <- colSums(is.na(returns_xts))
print(missing_summary)

# Display head of asset log returns
head(returns_xts) %>%
  kbl(caption = "First 6 Days of Asset Daily Log Returns (2020)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

# Exploratory Financial Analysis

```{r exploratory-stats}
# Summary statistics table
table.Stats(asset_returns_xts) %>%
  kbl(caption = "Descriptive Statistics for Portfolio Assets (2020--2025)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Correlation Structure

```{r correlation-matrix}
# Compute correlation matrix
cor_matrix <- cor(asset_returns_xts, use = "pairwise.complete.obs")

# Plot correlation heatmap
corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.col = "black",
  number.cex = 0.8,
  title = "Asset Return Correlation Matrix (Group 3 Universe)",
  mar = c(0, 0, 2, 0)
)
```

# Portfolio Construction (Equal-Weight vs Rebalanced)

```{r portfolio-construction}
# Placeholder for Portfolio Construction logic
```

# Portfolio Optimization (Efficient Frontier, GMV, Max Sharpe)

```{r portfolio-optimization}
# Placeholder for Portfolio Optimization routines
```

# Portfolio Risk Assessment (VaR, CVaR, Monte Carlo, Stress Testing)

```{r portfolio-risk}
# Placeholder for Portfolio Risk Assessment routines
```

# Executive Investment Recommendation

```{r executive-recommendation}
# Placeholder for final quantitative recommendation summary
```
