---
title: "Portfolio Management and Risk Consultation"
subtitle: "FINSTAD Case 2 -- Group 3 (Charlie)"
author:
  - "CRUZ, Ricardo Miguel Iñigo"
  - "GALEDO, Enrique Lorenzo Hermoso"
  - "SEBALLOS, Josiah Dweyn Panganiban"
  - "SEECHUNG, Camille Castro"
  - "SISON, Aaron Joshua Estacio"
date: "August 17, 2026"
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 3
    number-sections: true
    colorlinks: true
    documentclass: article
    classoption: 12pt
    geometry:
      - margin=1in
    fontsize: 11pt
    code-block-font-size: small
    highlight-style: arrow
    keep-tex: true
include-in-header:
  - text: |
      \usepackage{fvextra}
      \fvset{breaklines=true,breakanywhere=true}
      \RecustomVerbatimEnvironment{Highlighting}{Verbatim}{breaklines=true,breakanywhere=true,commandchars=\\\{\}}
      \let\verbatim\Verbatim
      \let\endverbatim\endVerbatim
      \usepackage{microtype}
      \usepackage{adjustbox}
      \usepackage{booktabs}
      \usepackage{tabularx}
      \usepackage{pdflscape}
      \usepackage{float}
      \usepackage{ragged2e}
      \usepackage{csquotes}
      \clubpenalty=10000
      \widowpenalty=10000
      \displaywidowpenalty=10000
      \sloppy
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 4.8,
  dpi = 300,
  fig.pos = "H"
)

# Safe kable helper for PDF and HTML rendering
safe_kable <- function(data, caption = NULL, digits = 4) {
  knitr::kable(data, caption = caption, digits = digits, booktabs = TRUE) %>%
    kableExtra::kable_styling(
      font_size = 8,
      latex_options = c("scale_down", "HOLD_position", "striped"),
      bootstrap_options = c("striped", "hover", "condensed")
    )
}
```

# Setup and Library Loading

Load the quantitative finance, portfolio analytics, and data visualization libraries required for the analysis.

```{r libraries}
# Core data manipulation and visualization
library(tidyverse)
library(lubridate)
library(scales)

# Time series and quantitative financial analysis
library(xts)
library(zoo)
library(PerformanceAnalytics)
library(PortfolioAnalytics)

# Solvers for portfolio optimization
library(ROI)
library(ROI.plugin.quadprog)
library(nloptr) 

# For Portfolio Risk Assessment
library(MASS) 

# Table formatting and visual aesthetics
library(corrplot)
library(knitr)
library(kableExtra)
```

# Data Ingestion and Preparation

Load the compiled master dataset containing historical daily close prices for Group 3's asset universe (TEL, MER, AEV, NVDA, META, BTC, SPY) and the risk-free rate benchmark over the sample period (January 2, 2020 -- January 2, 2026).

```{r data-ingestion}
# Define path to master dataset
data_path <- file.path("..", "data", "master_dataset.csv")

if (!file.exists(data_path)) {
  data_path <- file.path("data", "master_dataset.csv")
}

raw_data <- read_csv(data_path, show_col_types = FALSE)

# Preview structure
glimpse(raw_data)
```

## Daily Log Return Computation

Compute daily log returns for all portfolio assets and the benchmark:

$$r_{i,t} = \ln\left(\frac{P_{i,t}}{P_{i,t-1}}\right)$$

```{r return-computation}
prices_df <- raw_data %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date)

# Extract asset price columns
asset_cols <- c("TEL", "MER", "AEV", "NVDA", "META", "BTC", "SPY")

# Convert matrix to xts time series object
prices_mat <- as.matrix(prices_df[, asset_cols])
prices_xts <- xts(prices_mat, order.by = prices_df[["Date"]])

returns_xts <- Return.calculate(prices_xts, method = "log")[-1, ]

# Separate portfolio assets from benchmark
portfolio_assets <- c("TEL", "MER", "AEV", "NVDA", "META", "BTC")
benchmark_asset <- "SPY"

asset_returns_xts <- returns_xts[, portfolio_assets]
benchmark_returns_xts <- returns_xts[, benchmark_asset]

# No-trade days carry the last price mark, so the resumption-day return
# captures the full move across the gap. Wealth indexes use exp(cumsum(r))
# because these are log returns; cumprod(1 + r) would misstate growth.
prices_carried <- na.locf(prices_xts)
returns_xts_carried <- Return.calculate(prices_carried, method = "log")[-1, ]
asset_returns_carried <- returns_xts_carried[, portfolio_assets]

# EDA series: carried returns with no-trade days set back to NA, so
# resumption-day returns (full gap moves) enter the moments, correlations,
# and distribution figures while no-trade days stay missing.
returns_xts_eda <- returns_xts_carried
returns_xts_eda[is.na(prices_xts[-1, ])] <- NA
asset_returns_eda <- returns_xts_eda[, portfolio_assets]

# Risk-free rate (convert annual T-Bill rate to daily log-equivalent rate)
rf_annual <- mean(prices_df[["RF_Rate"]], na.rm = TRUE)
rf_daily <- log(1 + rf_annual) / 252

cat("Dataset span:", format(min(index(returns_xts))), "to", format(max(index(returns_xts))), "\n")
cat("Total trading days:", nrow(returns_xts), "\n")
cat("Annualized Risk-Free Rate:", percent(rf_annual, accuracy = 0.01), "\n")
cat("Daily Log Risk-Free Rate:", percent(rf_daily, accuracy = 0.0001), "\n")
```

## Data Validation and Summary Check

Verify data integrity, check for missing values, and print head records of returns.

```{r}
#| label: tbl-data-validation
#| tbl-cap: "First 6 Days of Asset Daily Log Returns (2020)"

# Missing days are expected for MER and AEV: Philippine holidays and genuine
# PSE closures on US trading days, plus one captured-data gap (2022-10-24).
# The ingestion script stores them as NA instead of carrying the last close,
# so no-trade days never become fake 0 returns.
missing_summary <- colSums(is.na(returns_xts))
notrade_summary <- colSums(is.na(returns_xts_eda))
cat("Missing return days per asset (raw gap-boundary convention):\n")
print(missing_summary)
cat("No-trade days per asset (gap-aware convention):\n")
print(notrade_summary)
if (any(missing_summary > 0)) {
  message("Missing return days present; verify no-trade day treatment in data acquisition")
}

# Format first 6 rows for table display
returns_head_df <- as.data.frame(head(returns_xts)) %>%
  rownames_to_column(var = "Date") %>%
  mutate(across(-Date, ~ sprintf("%.4f", .x)))

kbl(returns_head_df, booktabs = TRUE, align = "c") %>%
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = c("striped", "hover"))
```

# Exploratory Financial Analysis

```{r}
#| label: tbl-exploratory-stats
#| tbl-cap: "Descriptive Statistics for Portfolio Assets (Jan 2, 2020 -- Jan 2, 2026)"

stats_mat <- table.Stats(asset_returns_eda)

# Convert to data frame for clean rendering
stats_df <- as.data.frame(stats_mat) %>%
  rownames_to_column(var = "Metric")

kbl(stats_df, booktabs = TRUE, digits = 4) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), bootstrap_options = c("striped", "hover"))
```

## Quantitative Design System (QVRS Theme & Colors)

```{r quant-theme}
# QVRS Economist Color Tokens
qvrs_colors <- c(
  "TEL"  = "#1F2E7A", # Chicago 30
  "MER"  = "#2E45B8", # Chicago 45
  "AEV"  = "#475ED1", # Chicago 55
  "NVDA" = "#1DC9A4", # Hong Kong 45
  "META" = "#F97A1F", # Singapore 55
  "BTC"  = "#E3120B", # Economist Red
  "SPY"  = "#595959"  # London 35
)

# Custom Economist/QVRS ggplot2 theme
theme_quant <- function() {
  theme_minimal() +
    theme(
      text = element_text(color = "#0D0D0D"),
      plot.title = element_text(face = "bold", size = 12, margin = margin(b = 4)),
      plot.subtitle = element_text(size = 9.5, color = "#333333", margin = margin(b = 8)),
      plot.caption = element_text(size = 8, color = "#595959", hjust = 0, margin = margin(t = 6)),
      axis.title = element_text(size = 9, face = "bold", color = "#1A1A1A"),
      axis.text = element_text(size = 8, color = "#333333"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#F2F2F2", linewidth = 0.3),
      legend.position = "bottom",
      legend.title = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_rect(fill = "#EBEDFA", color = NA),
      strip.text = element_text(face = "bold", size = 9, color = "#141F52")
    )
}
```

## Historical Price & Return Trajectories

```{r}
#| label: fig-ts-returns
#| fig.cap: "Daily Log Return Trajectories Across Portfolio Universe (Jan 2020 -- Jan 2026)"
#| fig.width: 8
#| fig.height: 6

# Reshape returns for tidy plotting
returns_tidy <- as.data.frame(returns_xts_eda) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Asset", values_to = "Return") %>%
  mutate(Asset = factor(Asset, levels = names(qvrs_colors)))

ggplot(returns_tidy, aes(x = Date, y = Return, color = Asset)) +
  geom_line(linewidth = 0.35, alpha = 0.85) +
  facet_wrap(~ Asset, ncol = 2, scales = "free_y") +
  scale_color_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_date(date_labels = "%Y", date_breaks = "2 years") +
  labs(
    title = "Daily Log Return Dynamics and Volatility Clustering",
    subtitle = "Sample Period: January 2, 2020 through January 2, 2026",
    x = NULL,
    y = "Daily Log Return"
  ) +
  theme_quant() +
  theme(legend.position = "none")
```

## Distributional Shape: Histograms and Empirical Density vs Normal Reference

```{r}
#| label: fig-density-hist
#| fig.cap: "Empirical Return Distributions vs Gaussian Reference Curves"
#| fig.width: 8
#| fig.height: 6

# Compute normal distribution parameters per asset for reference curves
norm_params <- returns_tidy %>%
  group_by(Asset) %>%
  summarise(
    mean = mean(Return, na.rm = TRUE),
    sd = sd(Return, na.rm = TRUE),
    .groups = "drop"
  )

# Grid of x values for theoretical normal density overlay
grid_df <- norm_params %>%
  group_by(Asset) %>%
  reframe(
    x = seq(mean - 4 * sd, mean + 4 * sd, length.out = 200),
    y = dnorm(x, mean = mean, sd = sd)
  )

ggplot(returns_tidy, aes(x = Return)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 60,
    fill = "#D6DBF5",
    color = "#1F2E7A",
    linewidth = 0.2,
    alpha = 0.7
  ) +
  geom_density(aes(color = "Empirical Density"), linewidth = 0.8) +
  geom_line(
    data = grid_df,
    aes(x = x, y = y, color = "Gaussian Normal"),
    linetype = "dashed",
    linewidth = 0.75
  ) +
  facet_wrap(~ Asset, scales = "free", ncol = 3) +
  scale_color_manual(
    values = c("Empirical Density" = "#141F52", "Gaussian Normal" = "#E3120B")
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Empirical Distribution Shape: Heavy Tails and Non-Normality",
    subtitle = "Comparison of empirical kernel density against theoretical normal distribution",
    x = "Daily Log Return",
    y = "Density"
  ) +
  theme_quant()
```

## Quantile-Quantile (Q-Q) Tail Diagnostics

```{r}
#| label: fig-qq-plots
#| fig.cap: "Normal Q-Q Diagnostics Demonstrating Heavy Tail Leptokurtosis"
#| fig.width: 8
#| fig.height: 5.5

ggplot(returns_tidy, aes(sample = Return)) +
  stat_qq_line(color = "#E3120B", linewidth = 0.7, linetype = "dashed") +
  stat_qq(color = "#1F2E7A", alpha = 0.5, size = 0.8) +
  facet_wrap(~ Asset, scales = "free", ncol = 3) +
  labs(
    title = "Quantile-Quantile Tail Departure from Normality",
    subtitle = "S-shaped curves confirm pronounced fat-tail behavior across asset returns",
    x = "Theoretical Gaussian Quantiles",
    y = "Sample Quantiles"
  ) +
  theme_quant()
```

## Dispersion of Daily Returns and Outlier Diagnostics

```{r}
#| label: fig-boxplots
#| fig.cap: "Return Dispersion and Extreme Outlier Boxplots"
#| fig.width: 8
#| fig.height: 4.5

ggplot(returns_tidy, aes(x = Asset, y = Return, fill = Asset)) +
  geom_boxplot(outlier.color = "#E3120B", outlier.size = 1, outlier.alpha = 0.5, linewidth = 0.4) +
  scale_fill_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Cross-Asset Return Dispersion and Extreme Tail Outliers",
    subtitle = "Comparison of median, interquartile range (IQR), and extreme shocks",
    x = "Asset",
    y = "Daily Log Return"
  ) +
  theme_quant() +
  theme(legend.position = "none")
```

## Cumulative Performance Comparison (Growth of $1.00)

```{r}
#| label: fig-cum-returns
#| fig.cap: "Cumulative Growth Trajectories of USD 1.00 Investment (Jan 2020 -- Jan 2026)"
#| fig.width: 8
#| fig.height: 4.8

cum_returns_df <- as.data.frame(exp(cumsum(returns_xts_carried))) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Asset", values_to = "Wealth") %>%
  mutate(Asset = factor(Asset, levels = names(qvrs_colors)))

ggplot(cum_returns_df, aes(x = Date, y = Wealth, color = Asset)) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = qvrs_colors) +
  scale_y_log10(labels = label_dollar(prefix = "USD ")) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  labs(
    title = "Cumulative Wealth Index Growth Trajectories (Log Scale)",
    subtitle = "Hypothetical USD 1.00 investment performance from January 2020 to January 2026",
    x = NULL,
    y = "Wealth Index (USD)"
  ) +
  theme_quant()
```

## Correlation Structure

```{r}
#| label: fig-correlation-matrix
#| fig.cap: "Asset Return Correlation Heatmap (Group 3 Universe)"

cor_matrix <- cor(returns_xts_eda, use = "pairwise.complete.obs")

# Render correlation heatmap
corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.col = "black",
  number.cex = 0.8,
  mar = c(0, 0, 1, 0)
)
```

# Portfolio Construction (Equal-Weight vs Monthly Rebalanced)

This section evaluates the performance of equal-weighted asset allocations under two distinct execution regimes: a static Buy-and-Hold strategy and a disciplined Monthly Rebalanced strategy. The 6-asset portfolio universe (TEL, MER, AEV, NVDA, META, BTC) starts with equal baseline weights ($w_i = 1/6 \approx 16.67\%$) on January 2, 2020 and is benchmarked against the SPDR S&P 500 ETF Trust (SPY).

```{r portfolio-construction-calc}
# Simple (arithmetic) returns: a weighted sum of log returns is not the
# portfolio return, so Return.portfolio and all portfolio metrics below
# operate on simple returns converted from the carried log series.
ret_smp <- exp(returns_xts_carried) - 1
asset_returns_smp <- ret_smp[, portfolio_assets]

# Equal weight allocation vector across the 6 portfolio assets
w_eq <- rep(1 / 6, 6)
names(w_eq) <- colnames(asset_returns_smp)

# Strategy 1: Buy-and-Hold Equal-Weighted Portfolio
port_bnh <- Return.portfolio(
  R = asset_returns_smp,
  weights = w_eq,
  verbose = TRUE
)
r_bnh <- port_bnh$returns
weights_bnh <- port_bnh$EOP.Weight

# Strategy 2: Monthly Rebalanced Equal-Weighted Portfolio
port_reb <- Return.portfolio(
  R = asset_returns_smp,
  weights = w_eq,
  rebalance_on = "months",
  verbose = TRUE
)
r_reb <- port_reb$returns
weights_reb <- port_reb$EOP.Weight

# Risk-free rate metrics (5.25% per annum)
rf_annual <- 0.0525
rf_daily <- log(1 + rf_annual) / 252

# Performance evaluation helper function
# cum = prod(1+r)-1, CAGR = (1+cum)^(252/n)-1, vol = sd(r)*sqrt(252)
eval_portfolio <- function(r_series, name) {
  cum_ret <- as.numeric(tail(cumprod(1 + r_series), 1)) - 1
  n_obs   <- length(r_series)
  ann_ret <- (1 + cum_ret)^(252 / n_obs) - 1
  ann_sd  <- sd(as.numeric(r_series)) * sqrt(252)
  sharpe  <- (ann_ret - rf_annual) / ann_sd
  max_dd  <- as.numeric(maxDrawdown(r_series))
  
  data.frame(
    Strategy = name,
    `Cumulative Return` = sprintf("%.2f%%", cum_ret * 100),
    `CAGR (Annualized)` = sprintf("%.2f%%", ann_ret * 100),
    `Annual Volatility` = sprintf("%.2f%%", ann_sd * 100),
    `Sharpe Ratio (Rf=5.25%)` = sprintf("%.4f", sharpe),
    `Max Drawdown` = sprintf("%.2f%%", max_dd * 100),
    check.names = FALSE
  )
}

perf_summary_df <- rbind(
  eval_portfolio(r_bnh, "Equal-Weight (Buy & Hold)"),
  eval_portfolio(r_reb, "Equal-Weight (Monthly Rebalanced)"),
  eval_portfolio(ret_smp[, "SPY"], "SPDR S&P 500 ETF (SPY Benchmark)")
)
```

## Comparative Performance Evaluation

```{r}
#| label: tbl-portfolio-performance
#| tbl-cap: "Portfolio Performance Summary Comparison (Jan 2, 2020 -- Jan 2, 2026)"

safe_kable(perf_summary_df, caption = "Portfolio Performance Summary Comparison (Jan 2, 2020 -- Jan 2, 2026)")
```

## Cumulative Portfolio Growth Comparison

```{r}
#| label: fig-portfolio-growth
#| fig.cap: "Cumulative Growth Trajectories of USD 1.00 Investment Across Strategies"
#| fig.width: 8
#| fig.height: 5

# Merge portfolio return series (all simple returns)
port_returns_combined <- cbind(r_bnh, r_reb, ret_smp[, "SPY"])
colnames(port_returns_combined) <- c("Buy_Hold", "Monthly_Rebalanced", "SPY_Benchmark")

port_growth_df <- as.data.frame(cumprod(1 + port_returns_combined)) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Strategy", values_to = "Wealth") %>%
  mutate(
    Strategy = case_when(
      Strategy == "Buy_Hold" ~ "Equal-Weight (Buy & Hold)",
      Strategy == "Monthly_Rebalanced" ~ "Equal-Weight (Monthly Rebalanced)",
      Strategy == "SPY_Benchmark" ~ "SPDR S&P 500 ETF (SPY)"
    ),
    Strategy = factor(
      Strategy,
      levels = c("Equal-Weight (Buy & Hold)", "Equal-Weight (Monthly Rebalanced)", "SPDR S&P 500 ETF (SPY)")
    )
  )

strat_colors <- c(
  "Equal-Weight (Buy & Hold)" = "#E3120B",        # Economist Red
  "Equal-Weight (Monthly Rebalanced)" = "#1F2E7A", # Chicago 30
  "SPDR S&P 500 ETF (SPY)" = "#595959"            # London 35
)

ggplot(port_growth_df, aes(x = Date, y = Wealth, color = Strategy, linetype = Strategy)) +
  geom_line(linewidth = 0.85) +
  scale_color_manual(values = strat_colors) +
  scale_linetype_manual(values = c("solid", "solid", "dashed")) +
  scale_y_log10(labels = label_dollar(prefix = "USD ")) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  labs(
    title = "Cumulative Wealth Growth Comparison (Log Scale)",
    subtitle = "Performance of USD 1.00 allocation from January 2020 through January 2026",
    x = NULL,
    y = "Wealth Index (USD)"
  ) +
  theme_quant()
```

## Portfolio Asset Weight Drift Analysis

```{r}
#| label: fig-rebalanced-weights
#| fig.cap: "Asset Weight Drift Over Time in the Un-Rebalanced Buy-and-Hold Portfolio"
#| fig.width: 8
#| fig.height: 5.2

# Reshape end-of-period weights for Buy & Hold strategy
weights_bnh_df <- as.data.frame(weights_bnh) %>%
  rownames_to_column(var = "Date") %>%
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(-Date, names_to = "Asset", values_to = "Weight") %>%
  mutate(Asset = factor(Asset, levels = names(qvrs_colors)))

ggplot(weights_bnh_df, aes(x = Date, y = Weight, fill = Asset)) +
  geom_area(alpha = 0.85, color = "#FFFFFF", linewidth = 0.1) +
  scale_fill_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  labs(
    title = "Asset Allocation Drift in Buy-and-Hold Execution",
    subtitle = "Evolution of asset weights driven by differential market performance (Jan 2020 -- Jan 2026)",
    x = NULL,
    y = "Portfolio Weight Share"
  ) +
  theme_quant()
```

# Portfolio Optimization (Efficient Frontier, GMV, Max Sharpe)

```{r portfolio-optimization}
# Annualize mean returns and covariance matrix from daily log returns
mu_ann    <- colMeans(asset_returns_xts) * 252
Sigma_ann <- cov(asset_returns_xts) * 252

n_assets    <- length(mu_ann)
asset_names <- names(mu_ann)

# Equal-weight starting point for the solver
w0 <- rep(1 / n_assets, n_assets)
```

```{r}
# --- Objective 1: portfolio variance (minimized for GMV and each Efficient Frontier point) ---
eval_f_var <- function(w) {
  as.numeric(t(w) %*% Sigma_ann %*% w)
}

# Gradient via finite differences (nloptr's built-in numerical gradient helper)
eval_grad_f_var <- function(w) {
  nl.grad(w, eval_f_var)
}

# --- Objective 2: negative Sharpe ratio (minimized to find the tangency / Max Sharpe portfolio) ---
eval_f_sharpe <- function(w) {
  port_ret <- sum(w * mu_ann)
  port_sd  <- sqrt(as.numeric(t(w) %*% Sigma_ann %*% w))
  -(port_ret - rf_annual) / port_sd  # negative since nloptr always minimizes
}

eval_grad_f_sharpe <- function(w) {
  nl.grad(w, eval_f_sharpe)
}

# --- Constraint: weights must sum to 1 (fully invested, no cash drag) ---
eval_g_eq_sum1 <- function(w) {
  sum(w) - 1
}

eval_jac_g_eq_sum1 <- function(w) {
  nl.jacobian(w, eval_g_eq_sum1)
}

# Solver settings: SLSQP as guided, since it's the R equivalent of Excel's GRG Nonlinear
opt_settings <- list(
  "algorithm" = "NLOPT_LD_SLSQP",
  "xtol_rel"  = 1.0e-8,
  "maxeval"   = 1000
)
```

```{r}
# --- Global Minimum Variance (GMV) Portfolio ---
gmv_solution <- nloptr(
  x0            = w0,
  eval_f        = eval_f_var,
  eval_grad_f   = eval_grad_f_var,
  lb            = rep(0, n_assets),   # long-only, no shorting
  ub            = rep(1, n_assets),   # no single asset above 100%
  eval_g_eq     = eval_g_eq_sum1,
  eval_jac_g_eq = eval_jac_g_eq_sum1,
  opts          = opt_settings
)

w_gmv <- gmv_solution$solution
names(w_gmv) <- asset_names

gmv_ret <- sum(w_gmv * mu_ann)
gmv_sd  <- sqrt(as.numeric(t(w_gmv) %*% Sigma_ann %*% w_gmv))
```

```{r}
# --- Maximum Sharpe Ratio (Tangency) Portfolio ---
sharpe_solution <- nloptr(
  x0            = w0,
  eval_f        = eval_f_sharpe,
  eval_grad_f   = eval_grad_f_sharpe,
  lb            = rep(0, n_assets),
  ub            = rep(1, n_assets),
  eval_g_eq     = eval_g_eq_sum1,
  eval_jac_g_eq = eval_jac_g_eq_sum1,
  opts          = opt_settings
)

w_sharpe <- sharpe_solution$solution
names(w_sharpe) <- asset_names

sharpe_ret <- sum(w_sharpe * mu_ann)
sharpe_sd  <- sqrt(as.numeric(t(w_sharpe) %*% Sigma_ann %*% w_sharpe))
sharpe_val <- (sharpe_ret - rf_annual) / sharpe_sd
```

```{r}
# --- Efficient Frontier: sweep target returns, minimize variance at each one ---
target_returns <- seq(min(mu_ann), max(mu_ann), length.out = 50)
frontier_list  <- vector("list", length(target_returns))

for (i in seq_along(target_returns)) {
  target <- target_returns[i]

  # Two equality constraints at this point: weights sum to 1, AND portfolio return hits target
  eval_g_eq_frontier <- function(w) {
    c(sum(w) - 1, sum(w * mu_ann) - target)
  }
  eval_jac_g_eq_frontier <- function(w) {
    nl.jacobian(w, eval_g_eq_frontier)
  }

  sol <- tryCatch(
    nloptr(
      x0            = w0,
      eval_f        = eval_f_var,
      eval_grad_f   = eval_grad_f_var,
      lb            = rep(0, n_assets),
      ub            = rep(1, n_assets),
      eval_g_eq     = eval_g_eq_frontier,
      eval_jac_g_eq = eval_jac_g_eq_frontier,
      opts          = opt_settings
    ),
    error = function(e) NULL
  )

  # Only keep the point if the solver actually converged
  if (!is.null(sol) && sol$status >= 0) {
    w_i <- sol$solution
    frontier_list[[i]] <- data.frame(
      Target_Return = target,
      Risk   = sqrt(as.numeric(t(w_i) %*% Sigma_ann %*% w_i)),
      Return = sum(w_i * mu_ann)
    )
  }
}

efficient_frontier_df <- bind_rows(frontier_list)
```

```{r}
#| label: fig-efficient-frontier
#| fig.cap: "Efficient Frontier with Global Minimum Variance and Maximum Sharpe Portfolios"
#| fig.width: 8
#| fig.height: 5

ggplot(efficient_frontier_df, aes(x = Risk, y = Return)) +
  geom_line(color = "#1F2E7A", linewidth = 0.9) +
  geom_point(
    data = data.frame(Risk = gmv_sd, Return = gmv_ret),
    aes(x = Risk, y = Return), color = "#E3120B", size = 3
  ) +
  geom_point(
    data = data.frame(Risk = sharpe_sd, Return = sharpe_ret),
    aes(x = Risk, y = Return), color = "#1DC9A4", size = 3
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1)) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Efficient Frontier: Risk-Return Tradeoff",
    subtitle = "Red point = Global Minimum Variance | Teal point = Maximum Sharpe Ratio",
    x = "Annualized Volatility (Risk)",
    y = "Annualized Expected Return"
  ) +
  theme_quant()
```

```{r}
#| label: tbl-optimal-weights
#| tbl-cap: "Optimal Portfolio Weights: GMV vs Maximum Sharpe"

optimal_weights_df <- data.frame(
  Asset             = asset_names,
  GMV_Weight        = sprintf("%.2f%%", w_gmv * 100),
  MaxSharpe_Weight  = sprintf("%.2f%%", w_sharpe * 100)
)

safe_kable(optimal_weights_df, caption = "Optimal Portfolio Weights: GMV vs Maximum Sharpe")
```

```{r}
#| label: fig-portfolio-allocation
#| fig.cap: "Optimal Portfolio Allocation by Asset (GMV vs Max Sharpe)"
#| fig.width: 8
#| fig.height: 4.5

weights_long_df <- data.frame(
  Asset     = rep(asset_names, 2),
  Portfolio = rep(c("GMV", "Max Sharpe"), each = n_assets),
  Weight    = c(w_gmv, w_sharpe)
)

ggplot(weights_long_df, aes(x = Asset, y = Weight, fill = Asset)) +
  geom_col() +
  facet_wrap(~ Portfolio) +
  scale_fill_manual(values = qvrs_colors) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Optimal Portfolio Allocation by Asset",
    subtitle = "Comparison between Global Minimum Variance and Maximum Sharpe portfolios",
    x = NULL, y = "Portfolio Weight"
   ) +
  theme_quant() +
  theme(legend.position = "none")
```


# Portfolio Risk Assessment (VaR, CVaR, Monte Carlo, Stress Testing)

```{r portfolio-risk}
# Apply GMV and Max Sharpe weights to historical returns (monthly rebalanced,
# consistent with the rebalancing approach used in Portfolio Construction)
port_returns_gmv <- Return.portfolio(
  R = asset_returns_xts,
  weights = w_gmv,
  rebalance_on = "months"
)

port_returns_sharpe <- Return.portfolio(
  R = asset_returns_xts,
  weights = w_sharpe,
  rebalance_on = "months"
)
```

```{r}
#| label: tbl-var-summary
#| tbl-cap: "Historical and Parametric VaR / CVaR Summary (Daily, 95% and 99% Confidence)"

# --- 1. Compute Individual VaR and ES Metrics ---
# Global Minimum Variance (GMV)
v_hist_gmv_95 <- abs(as.numeric(VaR(port_returns_gmv, p = 0.95, method = "historical")))
v_hist_gmv_99 <- abs(as.numeric(VaR(port_returns_gmv, p = 0.99, method = "historical")))
v_para_gmv_95 <- abs(as.numeric(VaR(port_returns_gmv, p = 0.95, method = "gaussian")))
v_para_gmv_99 <- abs(as.numeric(VaR(port_returns_gmv, p = 0.99, method = "gaussian")))

e_hist_gmv_95 <- abs(as.numeric(ES(port_returns_gmv, p = 0.95, method = "historical")))
e_hist_gmv_99 <- abs(as.numeric(ES(port_returns_gmv, p = 0.99, method = "historical")))
e_para_gmv_95 <- abs(as.numeric(ES(port_returns_gmv, p = 0.95, method = "gaussian")))
e_para_gmv_99 <- abs(as.numeric(ES(port_returns_gmv, p = 0.99, method = "gaussian")))

# Max Sharpe
v_hist_shp_95 <- abs(as.numeric(VaR(port_returns_sharpe, p = 0.95, method = "historical")))
v_hist_shp_99 <- abs(as.numeric(VaR(port_returns_sharpe, p = 0.99, method = "historical")))
v_para_shp_95 <- abs(as.numeric(VaR(port_returns_sharpe, p = 0.95, method = "gaussian")))
v_para_shp_99 <- abs(as.numeric(VaR(port_returns_sharpe, p = 0.99, method = "gaussian")))

e_hist_shp_95 <- abs(as.numeric(ES(port_returns_sharpe, p = 0.95, method = "historical")))
e_hist_shp_99 <- abs(as.numeric(ES(port_returns_sharpe, p = 0.99, method = "historical")))
e_para_shp_95 <- abs(as.numeric(ES(port_returns_sharpe, p = 0.95, method = "gaussian")))
e_para_shp_99 <- abs(as.numeric(ES(port_returns_sharpe, p = 0.99, method = "gaussian")))

# --- 2. Construct Data Frame ---
risk_summary_df <- data.frame(
  Portfolio = rep(c("GMV", "Max Sharpe"), each = 4),
  Metric    = rep(c("Historical VaR", "Parametric VaR", "Historical ES (CVaR)", "Parametric ES (CVaR)"), 2),
  `95% CI`  = sprintf("%.2f%%", c(
    v_hist_gmv_95, v_para_gmv_95, e_hist_gmv_95, e_para_gmv_95,
    v_hist_shp_95, v_para_shp_95, e_hist_shp_95, e_para_shp_95
  ) * 100),
  `99% CI`  = sprintf("%.2f%%", c(
    v_hist_gmv_99, v_para_gmv_99, e_hist_gmv_99, e_para_gmv_99,
    v_hist_shp_99, v_para_shp_99, e_hist_shp_99, e_para_shp_99
  ) * 100),
  check.names = FALSE
)

# --- 3. Render Table (styled consistently with the rest of the report) ---
safe_kable(risk_summary_df, caption = "Historical and Parametric VaR / CVaR Summary (Daily, 95% and 99% Confidence)")
```

```{r}
# Simulate 10,000 one-day return scenarios using the annualized mean/covariance
# structure, converted back to daily, drawn from a multivariate normal distribution
set.seed(123)
n_sims <- 10000

mu_daily    <- mu_ann / 252
Sigma_daily <- Sigma_ann / 252

sim_asset_returns <- mvrnorm(n = n_sims, mu = mu_daily, Sigma = Sigma_daily)
colnames(sim_asset_returns) <- asset_names

# Apply portfolio weights to each simulated draw
sim_port_returns_gmv    <- as.numeric(sim_asset_returns %*% w_gmv)
sim_port_returns_sharpe <- as.numeric(sim_asset_returns %*% w_sharpe)

# Monte Carlo VaR and ES (empirical quantile of the simulated distribution)
mc_var_gmv_95    <- quantile(sim_port_returns_gmv, 0.05)
mc_var_sharpe_95 <- quantile(sim_port_returns_sharpe, 0.05)

mc_es_gmv_95    <- mean(sim_port_returns_gmv[sim_port_returns_gmv <= mc_var_gmv_95])
mc_es_sharpe_95 <- mean(sim_port_returns_sharpe[sim_port_returns_sharpe <= mc_var_sharpe_95])

mc_summary_df <- data.frame(
  Portfolio = c("GMV", "Max Sharpe"),
  `MC VaR (95%)` = sprintf("%.2f%%", c(abs(mc_var_gmv_95), abs(mc_var_sharpe_95)) * 100),
  `MC ES (95%)`  = sprintf("%.2f%%", c(abs(mc_es_gmv_95), abs(mc_es_sharpe_95)) * 100),
  check.names = FALSE
)
```

```{r}
#| label: tbl-mc-summary
#| tbl-cap: "Monte Carlo Simulated VaR and Expected Shortfall (10,000 Simulations)"

safe_kable(mc_summary_df, caption = "Monte Carlo Simulated VaR and Expected Shortfall (10,000 Simulations)")
```

```{r}
# --- Historical Stress Scenario Windows ---
# Define real calendar date ranges for each crisis, based on documented market events.
# These slice directly into your actual return series instead of guessing shock magnitudes.

covid_crash_window     <- "2020-02-19/2020-03-23"  # Global equity market top to COVID-19 bottom
crypto_winter_window   <- "2021-11-10/2022-11-21"  # BTC all-time high to FTX-collapse low
fed_hike_window        <- "2022-01-01/2022-10-31"  # Aggressive Fed tightening cycle (2022)
tech_selloff_window    <- "2022-01-01/2022-06-30"  # Nasdaq-heavy growth/tech correction (H1 2022)

# Helper: compute cumulative (peak-to-trough style) return per asset over a given window
calc_window_shock <- function(returns_xts, date_range) {
  window_returns <- returns_xts[date_range]
  # Cumulative return over the window: compounding daily log returns
  sapply(window_returns, function(col) exp(sum(col, na.rm = TRUE)) - 1)
}

# Compute actual shocks per asset for each historical scenario
shock_covid   <- calc_window_shock(asset_returns_xts, covid_crash_window)
shock_crypto  <- calc_window_shock(asset_returns_xts, crypto_winter_window)
shock_fed     <- calc_window_shock(asset_returns_xts, fed_hike_window)
shock_tech    <- calc_window_shock(asset_returns_xts, tech_selloff_window)
```

```{r}
# --- Hypothetical Black Swan ---
# This one is intentionally NOT historical — stress testing frameworks (e.g. Basel,
# CCAR-style) standardly include a hypothetical scenario worse than anything observed,
# to test tail risk beyond the historical sample. Here it's constructed as a uniform
# amplification of the worst historical single-scenario shock per asset, clearly
# labeled as a hypothetical assumption in the report (not fabricated historical data).
worst_observed <- pmin(shock_covid, shock_crypto, shock_fed, shock_tech)
shock_black_swan <- worst_observed * 1.5  # 50% more severe than worst historical case on record
```

```{r}
# --- Assemble the stress scenario table from real computed values ---
stress_scenarios <- tibble(
  Scenario = c("2020 COVID Crash", "2022 Crypto Winter", "Fed Rate Hike Shock",
               "Tech Sector Selloff", "Hypothetical Black Swan"),
  TEL  = c(shock_covid["TEL"],  shock_crypto["TEL"],  shock_fed["TEL"],  shock_tech["TEL"],  shock_black_swan["TEL"]),
  MER  = c(shock_covid["MER"],  shock_crypto["MER"],  shock_fed["MER"],  shock_tech["MER"],  shock_black_swan["MER"]),
  AEV  = c(shock_covid["AEV"],  shock_crypto["AEV"],  shock_fed["AEV"],  shock_tech["AEV"],  shock_black_swan["AEV"]),
  NVDA = c(shock_covid["NVDA"], shock_crypto["NVDA"], shock_fed["NVDA"], shock_tech["NVDA"], shock_black_swan["NVDA"]),
  META = c(shock_covid["META"], shock_crypto["META"], shock_fed["META"], shock_tech["META"], shock_black_swan["META"]),
  BTC  = c(shock_covid["BTC"],  shock_crypto["BTC"],  shock_fed["BTC"],  shock_tech["BTC"],  shock_black_swan["BTC"])
)
```

```{r}
#| label: fig-mc-distribution
#| fig.cap: "Historical vs Monte Carlo Simulated Return Distribution (Max Sharpe Portfolio)"
#| fig.width: 8
#| fig.height: 5

dist_compare_df <- bind_rows(
  data.frame(Return = as.numeric(port_returns_sharpe), Source = "Historical"),
  data.frame(Return = sim_port_returns_sharpe, Source = "Monte Carlo")
)

ggplot(dist_compare_df, aes(x = Return, fill = Source, color = Source)) +
  geom_density(alpha = 0.35, linewidth = 0.7) +
  scale_fill_manual(values = c("Historical" = "#1F2E7A", "Monte Carlo" = "#E3120B")) +
  scale_color_manual(values = c("Historical" = "#1F2E7A", "Monte Carlo" = "#E3120B")) +
  scale_x_continuous(labels = label_percent(accuracy = 1)) +
  labs(
    title = "Historical vs Monte Carlo Simulated Return Distribution",
    subtitle = "Max Sharpe Portfolio: comparing observed daily returns to 10,000 simulated draws",
    x = "Daily Portfolio Return",
    y = "Density"
  ) +
  theme_quant()
```

```{r}
# Historical Stress Scenario Windows (short, non-overlapping, event-anchored)
covid_crash_window     <- "2020-02-19/2020-03-23"  # Global equity market top to COVID-19 bottom
tech_selloff_window    <- "2021-11-19/2022-01-27"  # Initial growth-stock de-rating, pre-Fed-hike
crypto_winter_window   <- "2022-05-05/2022-05-13"  # Terra/Luna collapse window
fed_hike_window        <- "2022-06-08/2022-06-16"  # CPI-triggered selloff into June FOMC 75bp hike

# Helper: compute cumulative (peak-to-trough style) return per asset over a given window
calc_window_shock <- function(returns_xts, date_range) {
  window_returns <- returns_xts[date_range]
  sapply(window_returns, function(col) exp(sum(col, na.rm = TRUE)) - 1)
}

# Compute actual shocks per asset for each historical scenario
shock_covid   <- calc_window_shock(asset_returns_xts, covid_crash_window)
shock_tech    <- calc_window_shock(asset_returns_xts, tech_selloff_window)
shock_crypto  <- calc_window_shock(asset_returns_xts, crypto_winter_window)
shock_fed     <- calc_window_shock(asset_returns_xts, fed_hike_window)

# Hypothetical Black Swan: 1.5x the worst-observed shock per asset across the 4 historical scenarios
worst_observed   <- pmin(shock_covid, shock_crypto, shock_fed, shock_tech)
shock_black_swan <- worst_observed * 1.5

# Rebuild stress_scenarios from REAL computed values (replaces the fabricated tribble)
stress_scenarios <- tibble(
  Scenario = c("2020 COVID Crash", "2022 Crypto Winter", "Fed Rate Hike Shock",
               "Tech Sector Selloff", "Hypothetical Black Swan"),
  TEL  = c(shock_covid["TEL"],  shock_crypto["TEL"],  shock_fed["TEL"],  shock_tech["TEL"],  shock_black_swan["TEL"]),
  MER  = c(shock_covid["MER"],  shock_crypto["MER"],  shock_fed["MER"],  shock_tech["MER"],  shock_black_swan["MER"]),
  AEV  = c(shock_covid["AEV"],  shock_crypto["AEV"],  shock_fed["AEV"],  shock_tech["AEV"],  shock_black_swan["AEV"]),
  NVDA = c(shock_covid["NVDA"], shock_crypto["NVDA"], shock_fed["NVDA"], shock_tech["NVDA"], shock_black_swan["NVDA"]),
  META = c(shock_covid["META"], shock_crypto["META"], shock_fed["META"], shock_tech["META"], shock_black_swan["META"]),
  BTC  = c(shock_covid["BTC"],  shock_crypto["BTC"],  shock_fed["BTC"],  shock_tech["BTC"],  shock_black_swan["BTC"])
)
```

```{r}
# Compute portfolio-level impact for each scenario under both GMV and Max Sharpe weights
stress_impact_df <- stress_scenarios %>%
  rowwise() %>%
  mutate(
    shock_vec        = list(c(TEL, MER, AEV, NVDA, META, BTC)),
    GMV_Impact       = sum(unlist(shock_vec) * w_gmv[c("TEL","MER","AEV","NVDA","META","BTC")]),
    MaxSharpe_Impact = sum(unlist(shock_vec) * w_sharpe[c("TEL","MER","AEV","NVDA","META","BTC")])
  ) %>%
  ungroup() %>%
  dplyr::select(Scenario, GMV_Impact, MaxSharpe_Impact)
```

TEL, MER, and AEV showed relatively limited daily price variation throughout large portions of the sample period, which is consistent with the zero-valued median and interquartile range reported in the descriptive statistics in Section 3. Consequently, the stress shock estimates for these three assets, particularly AEV, may not fully reflect their actual responses during historical crisis periods. This data limitation should therefore be considered when interpreting the results.
```{r}
#| label: tbl-stress-summary
#| tbl-cap: "Portfolio Impact Under Stress Scenarios"

stress_summary_df <- stress_impact_df %>%
  mutate(
    GMV_Impact = sprintf("%.2f%%", GMV_Impact * 100),
    MaxSharpe_Impact = sprintf("%.2f%%", MaxSharpe_Impact * 100)
  ) %>%
  rename(`GMV Portfolio Impact` = GMV_Impact, `Max Sharpe Portfolio Impact` = MaxSharpe_Impact)

safe_kable(stress_summary_df, caption = "Portfolio Impact Under Stress Scenarios")
```

```{r}
#| label: fig-stress-test
#| fig.cap: "Portfolio Value Impact Across Stress Scenarios"
#| fig.width: 8
#| fig.height: 5

stress_long_df <- stress_impact_df %>%
  pivot_longer(cols = c(GMV_Impact, MaxSharpe_Impact), names_to = "Portfolio", values_to = "Impact") %>%
  mutate(Portfolio = ifelse(Portfolio == "GMV_Impact", "GMV", "Max Sharpe"))

ggplot(stress_long_df, aes(x = reorder(Scenario, Impact), y = Impact, fill = Portfolio)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = c("GMV" = "#1F2E7A", "Max Sharpe" = "#E3120B")) +
  labs(
    title = "Portfolio Impact Under Stress Scenarios",
    subtitle = "Simulated instantaneous shock to portfolio value, GMV vs Max Sharpe weighting",
    x = NULL,
    y = "Portfolio Return Impact"
  ) +
  theme_quant()
```

# Executive Investment Recommendation

```{r executive-recommendation}
# Placeholder for final quantitative recommendation summary
```
